---
# IMPORTANT: Change settings here, but DO NOT change the spacing AT ALL.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "22894551 - Question 2 - Play that Song"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Charisa Amorie Geyer"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Cape Town, South Africa" # First Author's Affiliation
Email1: "22894551\\@sun.ac.za" # First Author's Email address
# 
# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE
# 
# # Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
# <!-- abstract: | -->
# <!--   Abstract to be written here. The abstract should not be too long and should provide the reader with a good understanding what you are writing about. Academic papers are not like novels where you keep the reader in suspense. To be effective in getting others to read your paper, be as open and concise about your findings here as possible. Ideally, upon reading your abstract, the reader should feel he / she must read your paper in entirety. -->
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')

if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
library(ggridges)
library(readr)
library(lubridate)
library(dplyr)
library(ggridges)
library(ggplot2)
library(viridis)
library(hrbrthemes)
list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))


# upload data
coldplay <- read_csv("Data/Coldplay_vs_Metallica/Coldplay.csv")
metallica <- read_csv("Data/Coldplay_vs_Metallica/metallica.csv")
spotify <- read_rds("Data/Coldplay_vs_Metallica/Broader_Spotify_Info.rds")
billboard_100 <- read_rds("Data/Coldplay_vs_Metallica/charts.rds")

#create the year cols

coldplay <- coldplay %>%
  mutate(year = year(as.Date(release_date)))

metallica <- metallica %>% 
    mutate(year = year(as.Date(release_date)))

billboard_100 <- billboard_100 %>% 
    mutate(year = year(as.Date(date)))

# Let's name the columns consistently & remove special characters in all dataframes
metallica <- metallica %>% rename(song = name) %>% mutate(artist = "Metallica") 
coldplay <- coldplay %>% rename(album = album_name, song = name) %>%  mutate(artist = "Coldplay")
spotify <- spotify %>% rename(song = name)
# billboard is fine

# remove_special characters
glimpse(metallica)

metallica <- remove_special(metallica, album)
coldplay <- remove_special(coldplay, album)

# filter only for studio albums!

metallica_studio <- metallica %>%
  filter(!grepl("live|Live|LIVE|Remastered|Remaster|Deluxe", album, ignore.case = TRUE))
coldplay_studio <- coldplay %>% 
    filter(!grepl("live|Live|LIVE|Remastered|Demo|Deluxe", album, ignore.case = TRUE))

metallica_live <- metallica %>% 
    filter(grepl("live|Live|LIVE", album, ignore.case = T))

coldplay_live <- coldplay %>% 
    filter(grepl("live|Live|LIVE", album, ignore.case = T))

# create a list of spotify songs that reached the top 100 on billboard
spotify_top100 <- top100(spotify)


# create one dataframe wit coldplay and metallica studio albums together
together_studio <- bind_rows(metallica_studio, coldplay_studio)

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction

This is an analysis of two iconic bands with >20-year careers. Firstly, Coldplay (Alternative/Pop Rock, debut: 2000), and secondly Metallica (Heavy Metal/Thrash, debut: 1981). Using Spotify data, we can compare their musical progression, audience engagement, and adaptation to industry trends, filtering only studio recordings for fair comparisons. 

\newpage
# Popularity by album
A box-and-whisker plot compares the popularity scores of each band's albums, revealing trends across their discographies.

## Coldplay
Coldplay has released a wide range of albums over the years. Coldplay's evolving sound over the years is reflected in the varying popularity of their studio and live albums.

### Overall
```{r}
albumpop_CP <- boxplotpopbyalbum(coldplay)
albumpop_CP
```
Coldplay is known for its live albums, with the live albums of "Live in Bueonos Aires" and "Live 2012" surpassing albums like "Love in Tokyo" and "Everyday Life" in popularity.  
### Live

```{r}
albumpop_CPL <- boxplotpopbyalbum(coldplay_live)
albumpop_CPL 
```
Live albums consistently rank highly, suggesting stronger fan engagement with concert performances.

\newpage
## Metallica
Metallica's enduring appeal is evident, with remastered editions dominating their most popular works.
### Overall

```{r}
albumpop_ML <- boxplotpopbyalbum(metallica)
albumpop_ML
```
Remastered albums claim 7 of the top 10 spots, highlighting their continued relevance and enhanced production value. 
### Studio

```{r}
albumpop_MLS <- boxplotpopbyalbum(metallica_studio)
albumpop_MLS
```
Classic releases maintain a strong presence, showing a loyal fanbase. Streaming services have made Metallica's music much more accessible. Seeing an steady increase in popularity over time is unsurprising.
\newpage

# Song popularity
Now let's look deeper into the popularity of the songs within each album. 
```{r fig.align = 'center', fig.pos = "H", out.width = "80%"}
songspop <- songsinalbum_pop(together_studio)
songspop
```
"Parachutes" and "A Rush of Blood to the Head" are identified as Coldplay's two most popular albums of all time. Within those two albums, all songs are popular. Compared to "Mylo Xyoto", where many songs are popular, but some are unpopular - bumping the album to a lower ranking overall. This shows the importance of song popularity in determining album popularity. 

\newpage

# Timetrends in music style
To look deeper into the longevity of these popular bands, we can consider the development of music styles over time. By reviewing the time trends of components of their music compared to that of Spotify (a proxy for global music trends), we can uncover whether the artists changed their style to meet global trends. Metallica, known as a loud band, remains louder than the global trend, up until about 2015, maintaining their signature style. 
```{r}
timetrend_ML_loudness <- timetrend_comp(df = metallica_studio, artist = "Metallica", variable = "loudness", min_year = 1990)
timetrend_ML_loudness
```
Coldplay starts following the global danceability trends, but returns to their calmer style over time. 
```{r}
timetrend_CP_dance <- timetrend_comp(df = coldplay_studio, artist = "Coldplay", variable = "danceability", min_year = 2000)
timetrend_CP_dance
```

\newpage
# Live vs Studio albums
We can see that energy trends are much lower in Live albums, for both Coldplay and Metallica, but especially for Metallica. 
```{r fig.align = 'center', fig.pos = "H", out.width = "80%"}
liveCP <- live_timetrend_comp(coldplay, coldplay_live, artist_name = "Coldplay Studio", compare_name = "Live Albums", variable = "energy")
liveCP

liveML <- live_timetrend_comp(metallica, metallica_live, artist_name = "Metallica Studio", compare_name = "Live Albums", variable = "energy")
liveML
```

# Conclusion 
Both Metallica and Coldplay demonstrate how iconic artists balance staying true to their roots while navigating evolving music trends. Ultimately, their sustained popularity suggests that success lies not in chasing trends, but in blending selective innovation with core artistic identity.
