---
# IMPORTANT: Change settings here, but DO NOT change the spacing AT ALL.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "22894551 - Question 5 - An Apple a Day: Hack your Health"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Charisa Amorie Geyer"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Cape Town, South Africa" # First Author's Affiliation
Email1: "22894551\\@sun.ac.za" # First Author's Email address
# 
# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE
# 
# # Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
  # <!-- abstract: | -->
# <!--   Abstract to be written here. The abstract should not be too long and should provide the reader with a good understanding what you are writing about. Academic papers are not like novels where you keep the reader in suspense. To be effective in getting others to read your paper, be as open and concise about your findings here as possible. Ideally, upon reading your abstract, the reader should feel he / she must read your paper in entirety. -->
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!

# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
library(readr)
library(tidyverse)
library(lubridate)
library(stringr)
library(readr)
library(ggridges)
library(ggplot2)
library(viridis)
library(hrbrthemes)
library(ggstream)
library(broom)
library(tidyfit)
library(fmxdat)

list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))
Health <- "./data/Health/HealthCare.csv" %>% read_csv()
Health_lpm <- thresholdBMR(Health)
```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Slide 1: Intro

Taking care of one's health is crucial because it impacts overall well-being, including physical, mental, and emotional health, which in turn affects daily life, productivity, and longevity. Good health allows individuals to function optimally, cope with stress, and enjoy a higher quality of life. 

# Slide 2: What is BMR? 
Basal metabolic rate is the speed at which your body performs basic metabolic functions. 
It the rate at which your body burns calories while at rest, performing essential functions like breathing and circulation. It represents the minimum number of calories needed to keep your body functioning at a basic level. Essentially, it's the energy your body uses just to stay alive when you're not actively doing anything. 

# Slide 3: BMR threshold
This threshold is calculated (differently for males and female). It consists of a metric similar to BMI, where you take a transformed value of your height, weight and activity levels to come to a factor. It is sometimes calculated to see how many calories one should consume to maintain your current weight. 
\newpage
# Slide 4: Determinants of Health
Source: @sprintmedical_bmr_2022
* Age -> older people have slower metabolisms
* Muscle mass -> muscles burn calories
* Body size -> having bigger organs requires more energy to keep you alive
* Gender -> Men and Women have different bodily functions and hormone levels
* Physical Activity -> strenous activity requires energy
* Diet -> macronutrient intake makes a difference, more than just calories

# Silde 5: Who is healthy
We can see here, that in this dataset of persons between 20 and 60, the activity levels indicate an unexpected result. Further, stress levels are also not following a clear trend. This is indicative of another underlying fcactor determining the health.  
```{r}
g <- stackedbar_BMR(Health_lpm, category_col = "Stress Level")
 h <- stackedbar_BMR(Health_lpm, category_col = "Physical Activity Level")
cowplot::plot_grid(g, h, ncol = 1)
```

\newpage

# Slide 6: LPM results
Men and women experience different effects. Remember that an LPM measures effects relative to a basline category, which in this case is low stress levels with excellent sleep and no activity. We can see clearly (as clearly as LPM results can be), that excellent sleep is always better, more moderate activity is acceptable and stress levels have varying effects.  

```{r}
coefplot <- LPM(Health_lpm)
coefplot
```

# What's next? 
* sleep more
* exercise less
* stop stressing about your stress levels!
This is of course a joke, and data can always be wrangled in a way to fit a nrrative. So the lesson to learn is: DON'T BELIEVE EVERYTHING THAT YOU SEE! Especially not if it comes from an amateur analyst! 

# Greetings
Thanks for listening and see you next time. 