---
# IMPORTANT: Change settings here, but DO NOT change the spacing AT ALL.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "22894551 - Question 1 - US Baby names"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Charisa Geyer"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Cape Town, South Africa" # First Author's Affiliation
Email1: "22894551\\@sun.ac.za" # First Author's Email address
# 
# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE
# 
# # Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  This report shows trends in US baby names. It looks at persistence and populatarity of baby names and its correlation to other popular US celebrities and Netflix/ TV shows over the decades. 
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')



rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)

list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

# upload the data
library(readr)
Baby_Names <- read_rds("./data/US_Baby_names/Baby_Names_By_US_State.rds")
Top_100_Billboard <- read_rds("./data/US_Baby_names/charts.rds")
HBO_titles <- read_rds("./data/US_Baby_names/HBO_titles.rds")
HBO_credits <- read_rds("./data/US_Baby_names/HBO_credits.rds")
```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}


# Plot 1: Popular names by decade

```{r, echo=FALSE}
library(tidyverse)
library(lubridate)
library(ggplot2)
# apply the top25_plot function
top25_girlsplot <- top25_plot(df = Baby_Names, xaxis_size = 10, xaxis_rows = 2, gender_filter = "F")
top25_girlsplot

top25_boysplot <- top25_plot(df = Baby_Names, xaxis_size = 10, xaxis_rows = 2, gender_filter = "M")
top25_boysplot

# since I want to do boys and girls separately,, I create boys df and girls df -. but the others aren't made like that so maybe not? 
```



```{r Figure1, warning =  FALSE, fig.align = 'center', fig.cap = "Caption Here \\label{Figure1}", fig.height = 3, fig.width = 6, dev = 'png'}
 
```

To make your graphs look extra nice in latex world, you could use Tikz device. Replace dev - 'png' with 'tikz' in the chunk below. Notice this makes the build time longer and produces extra tex files - so if you are comfortable with this, set your device to Tikz and try it out:

```{r, echo=FALSE}
library(stringr)

HBO_credits_MF <- add_gender(HBO_credits, name = Name)
Billboard_MF <- add_gender(Top_100_Billboard, name = artist)

```

\newpage

# References {-}

<div id="refs"></div>


# Appendix {-}

## Appendix A {-}

Some appendix information here

## Appendix B {-}

