---
# IMPORTANT: Change settings here, but DO NOT change the spacing AT ALL.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "22894551 - Question 1 - I was named after WHO?"
#subtitle: "This will appear as Right Header"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Charisa Geyer"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Stellenbosch University, Cape Town, South Africa" # First Author's Affiliation
Email1: "22894551\\@sun.ac.za" # First Author's Email address
# 
# Author2: "John Smith"
# Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "John\\@gmail.com"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.
# 
# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"
# 
# CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE
# 
# # Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: TRUE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
# header-includes:
#    - \usepackage{colortbl} # Add additional packages here.

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: |
  This report shows trends in US baby names. It looks at persistence and populatarity of baby names and its correlation to other popular US celebrities and Netflix/ TV shows over the decades. 
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')


rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
if(!require("tidyverse")) install.packages("tidyverse")
library(tidyverse)
library(stringr)
library(readr)
library(lubridate)
library(ggplot2)

list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

# upload the data

Baby_Names <- read_rds("./data/US_Baby_names/Baby_Names_By_US_State.rds")
Top_100_Billboard <- read_rds("./data/US_Baby_names/charts.rds")
HBO_titles <- read_rds("./data/US_Baby_names/HBO_titles.rds")
HBO_credits <- read_rds("./data/US_Baby_names/HBO_credits.rds")

#Join the HBO titles and credits
HBO_joined <- HBO_credits %>%
  left_join(HBO_titles, by = "id")

#tidy up the column names for ease of use: 
Baby_Names <- Baby_Names %>%
  rename(name = Name) %>% rename(year = Year)

# add gender & rename name columns

Top_100_Billboard <- Top_100_Billboard %>% rename(name = artist) #the dates is in numeric here, can add year col
Billboard_MF  <- add_gender(df = Top_100_Billboard)

#Hbo artists
HBO_MF <- add_gender(HBO_joined)
#now apply to character
HBO_characters_MF <- HBO_MF %>%  rename(actor_name = name) %>% rename(name = character) %>% rename(year = release_year)

#remove unnecessary dataframes:
rm(Top_100_Billboard, HBO_credits, HBO_joined, HBO_titles)


```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}
The Toy Store wants to know if there are trends in naming conventions. Parents get influenced by the media, tradition and other unknown factors. If trends are persistent, the Toy Store can rely on trends to cater their toy names to. 
\newpage

# Plot 1: Most popular names over the years
We observe the popularity of names for boys and girls over time. By looking at the colour-coordinated bubbles, we observe persistence in the naming of very popular names like "Mary" and "James" - the two most popular names overall. 
```{r, fig.align= 'center'}

# apply the top25_plot function
top25_girlsplot <- top25_plot(df = Baby_Names, xaxis_size = 20, xaxis_rows = 2, gender_filter = "F")

top25_boysplot <- top25_plot(df = Baby_Names, xaxis_size = 10, xaxis_rows = 2, gender_filter = "M")

cowplot::plot_grid(top25_boysplot, top25_girlsplot, nrow = 1)
```
\newpage

# Plot 2: Increase in name popularity year-on-year, per decade for Girls

These plots show how year-on-year popularity of names are distributed. The 3 names that had the highest uptake per decade is shown in red. This can be compared to releases of music. movies where the name is used. 
In early days, parents really did not look further than the first page of the "Baby Name Book". 
```{r}
q <- top25_increaseplot(Baby_Names)
q
```
\newpage

# Plot 3: Increase in name popularity year-on-year, per decade for boys

These plots show how year-on-year popularity of names are distributed. The 3 names that had the highest uptake per decade is shown in red. This can be compared to releases of music. movies where the name is used. 
Parents still looked at A-names, and keep on naming their boys names from the Bible. What is not seen here is the persistence of the name "James" Since it remains popular year-on-year, the increase may be less pertinent. 
```{r}
b <- top25_increaseplot(Baby_Names, gender_filter = "M", xaxis_rows =2, xaxis_size = 10)
b
#cowplot::plot_grid(q, b, nrow = 1)

```


\newpage

# Plot 4: Spearman correlation

```{r fig.spearman, fig.cap = "Figure: Spearman correlation for naming persistence over time"}
spearman_full(Baby_Names)
```

We can observe the Spearman rank correlations, portraying naming persistence of the Top 25girls and boys names in Figure \@ref(fig:spearman). In the early years, naming persistence was very high, with a correlation close to 1. Naming persistence decreased severely in the 1960s and stayed lower. For girls, it also decreased, but not as dramatically. Between the 1970s and 1990s, naming patterns in the United States began to shift noticeably, particularly for girls. The rise of television, pop music, and mass media exposed parents to a wider range of influences. Parents were now getting inspiration from soap opera characters to chart-topping musicians — which led to short-lived spikes in name popularity [@abramson2003TVhistory]. Names like Katina (influenced by the 1974 soap "Where the Heart Is") illustrate how easily cultural moments could shape naming choices. This era marked a move away from traditional, family-based naming toward more trend-driven and individualistic preferences, contributing to a clear decline in naming persistence as measured by year-over-year Spearman correlations. This leads us to looking into the correlation between the Top-100 Billoard artist names and HBO character and actor names, with popular baby names in the US. 

\newpage

# Comparison with Musicians, Song Titles and Prominent Figures

Popular names such as Dorothy and Elizabeth are likely to be based on political figures, Queen Elizabeth and Dorothy is strongly associated with the character Dorothy Gale from "The Wizard of Oz," further solidifying its place in popular culture. The 1939 film adaptation starring Judy Garland cemented the name's popularity. Barbara Stanwyck was establishing herself as a prominent film actress, appearing in popular movies like "Stella Dallas".

The most popular boys' names remain those of the 12 disciples in the Bible. 

I would advise the Toy Store to reconsider relying on celebrity names as a predictor of baby names, but rather rely on most common names to date. 

```{r results='asis', message=FALSE, warning=FALSE}

library(xtable)

# Create the data
name_table_data <- data.frame(
  Name = c("Donna", "Dorothy", "Diane", "Barbara", "Elizabeth"),
  Type = c("Song name", "Character", "Song", "Actress", "Public figure"),
  Example = c(
    '"Donna" - Ritchie Valens',
    'Dorothy Gale - The Wizard of Oz',
    '"Jack and Diane" - John Mellencamp',
    'Barbra Streisand', 
    'Queen Elizabeth'
  )
)

# Create xtable
name_xtable <- xtable(name_table_data, 
                     caption = "Names in Music, Film and Culture \\label{tab:names}")

# Print with formatting
print(name_xtable,
      include.rownames = FALSE,
      floating = TRUE,
      table.placement = 'H',
      caption.placement = 'top',
      comment = FALSE)
```

# References {-}

<div id="refs"></div>


